---
title: "Assignment02"
author: "Dingyuan Gu"
date: "2024-09-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




## Load packages


```{r load}
library(tidyverse)
library(Stat2Data)
data("Hawks")
head(Hawks)
```


# 1.Data Wrangling

## 1.1 Select and filter

### 1.1 (Q1)
```{r}
hSF<-Hawks %>% filter(Species== "RT" & Weight>=1000) %>% select(Wing, Weight, Tail)
print(hSF)
```
### 1.1 (Q2)

#### How many variables does the data frame hSF have?

hSF have 3 variables.


#### What would you say to communicate this information to a Machine Learning practitioner?

This data frame has 3 features, namely Wing, Weight, and Tail.

#### How many examples/observations/cases does the data frame hSF have?

398

## 1.2 The arrange function

### 1.2 (Q1)
```{R}
hSF %>% arrange(Wing) %>% head(5)
```
## 1.3 Join and rename functions

### 1.3 (Q1)

```{r}
hawkSpeciesNameCodes <- data.frame(
  species_code = c("CH", "RT", "SS"),
  species_name_full = c("Cooper's", "Red-tailed", "Sharp-shinned")
)
print(hawkSpeciesNameCodes)
```

### 1.3 (Q2)

```{r}
hawksFullName <- Hawks %>% select(Month, Day, Year, CaptureTime, ReleaseTime, BandNumber, Species) %>% left_join(hawkSpeciesNameCodes %>% rename(Species=species_code)) %>% left_join(Hawks)  %>% select(-Species) %>% rename(Species=species_name_full)
print(hawksFullName)
```

### 1.3 (Q3)

```{r}
print(hawksFullName %>% select(Species, Wing, Weight) %>% head(7))
```

## 1.4 The mutate function

### 1.4 (Q1)

```{r}
hawksWithBMI<- Hawks %>% mutate(bird_BMI=1000* Weight/(Wing^2)) %>% select(Species, bird_BMI) %>% arrange(desc(bird_BMI))
print(hawksWithBMI)
```

## 1.5 Summarize and group-by functions

### 1.5 (Q1)

```{r}
hawksFullName %>% group_by(Species) %>% summarise(num_rows=n(), mn_wing=mean(Wing, na.rm=TRUE), nd_wing=median(Wing, na.rm=TRUE), , t_mn_wing=mean(Wing, trim=0.1, na.rm = TRUE), b_wt_ratio=max(Wing/Tail, na.rm=TRUE))
```

### 1.5 (Q1)

```{r}
Hawks %>%
  group_by(Species) %>%
  select(Wing, Weight, Culmen, Hallux, Tail, StandardTail, Tarsus, Crop) %>%
  summarise(across(everything(), ~ sum(is.na(.x))))
```












