---
title: "Assignment03"
author: "levitt"
date: "2024-09-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Load packages
```{r}
library(tidyverse)
library(Stat2Data)
data("Hawks")
```


## 1. Visualisation

### 1 (Q1)

```{r}
hawksSmall<-drop_na(select(Hawks,Age,Day,Month,Year,CaptureTime,Species,Wing,Weight,Tail))
dim(hawksSmall)
head(hawksSmall,5)
```

### 1 (Q2)

```{r}
univar_plot1<-ggplot(data=hawksSmall,aes(x=Weight))+xlab("Weight")
univar_plot1+geom_histogram(binwidth = 10)+ylab("Count")
```

### 1 (Q3)

```{r}
univar_plot2<-ggplot(data=hawksSmall,aes(x=Tail))+xlab("Tail length")
univar_plot2+geom_histogram(binwidth =1)+ylab("Count")
univar_plot2+geom_density(adjust=0.5)
univar_plot2+geom_density(adjust=2)
```

### 1 (Q4)

```{r}
ggplot(data=hawksSmall, aes(x=Tail,color=Species))+ geom_density(adjust=2)+ theme_bw() + xlab("Tail length") + ylab("Count")
```

### 1 (Q5)

```{r}
ggplot(data=hawksSmall, aes(x=Tail,y=Species,fill=Species))+ geom_violin()+ theme_bw() + xlab("Tail (mm)")+ ylab("Density")
```

### 1 (Q6)

```{r}
plot3<- ggplot(data=hawksSmall, aes(x=Tail, y=Weight)) + xlab("Tail (mm)") + ylab("Weight (mm)")
plot3+geom_point(aes(color=Species, shape=Species))
```

#### Aesthetics

Tail -> horizontal position.
Weight -> vertical position.
Species -> color.
Species -> shape.

#### glyphs

CH -> circle
RT -> triangle
SS -> square

#### visual cues

The plot uses **color, shape, and position** as primary visual cues to represent different species and their corresponding data points.

### 1 (Q7)

```{r}
plot3+geom_point(aes(color = Species))+facet_wrap(~Species, scales = "free_x")+geom_smooth(method ="lm", aes(color=Species))
```








#### visual cues

The visual cues in this plot include faceting, color, position, and trend lines with confidence intervals to compare relationships between tail length and weight for each species.



#### relationship

There is a general positive relationship between the hawks' tail lengths and their weights. However, the strength of this relationship varies among species, with the RT species showing the most consistent correlation. The relationship for SS is less pronounced due to the smaller range of values.


### 1 (Q8)
```{r}
hawksSmall%>% select(Weight, Tail) %>% arrange(desc(Weight))%>%head(1)
```




```{r}
plot3+geom_point(aes(color=Species))+ 
  geom_curve(x=196, xend=196, y=1800, yend=2030, arrow=arrow(length=unit(0.1, 'cm')), curvature=0.5) + 
  geom_text(x=196, y=1800, label="heaviest hawk")
```


## 2. Finite probability spaces

### 2.1 (Q1)

$$
P=\left ( 0.3  \right ) ^{z}\times \left ( 0.7  \right ) ^{22-z}
$$
### 2.1 (Q2)

$$
P(X=z)=\binom{n}{z} (0.3)^{z}\times(0.7)^{n-z}, n=22
$$
```{r}
prob_red_spheres <- function(z) {
  n <- 22
  p <- 3/10  
  
  prob <- choose(n, z) * (p^z) * ((1 - p)^(n - z))
  
  return(prob)
}

prob_red_spheres(10)
```

### 2.1 (Q3)

```{r}
prob_by_num_reds<-data.frame(
  num_reds=c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22)
)
prob_by_num_reds$prob <- choose(22, prob_by_num_reds$num_reds) * (0.3^prob_by_num_reds$num_reds) * ((0.7)^(22 - prob_by_num_reds$num_reds))
head(prob_by_num_reds,3)
```


### 2.1 (Q4)

```{r}
plot4<-ggplot(data=prob_by_num_reds,aes(x=num_reds,y=prob))+geom_line()
plot4+xlab("Number of reds")+ylab("Probability")
```

### 2.1 (Q5)

```{r}
itermap <- function(.x, .f) {
 result <- list()
 for (item in .x) {
 result <- c(result, list(.f(item)))
 }
 return(result)
}


num_trials<-1000 # set the number of trials
set.seed(0) # set the random seed
sampling_with_replacement_simulation<-data.frame(trial=1:num_trials) %>% mutate(sample_balls = itermap(.x=trial, function(x){sample(10,22, replace = TRUE)}),
    num_reds = map_int(sample_balls, function(sample) sum(sample <= 3)))
print(sampling_with_replacement_simulation)
```

### 2.1 (Q6)

```{r}
itermap_dbl <- function(.x, .f) {
 result <- numeric(length(.x))
 for (i in 1:length(.x)) {
 result[i] <- .f(.x[[i]])
 }
 return(result)
}

num_reds_in_simulation<-sampling_with_replacement_simulation %>% pull(num_reds) 
# we extract a vector corresponding to the number of reds in each trial
prob_by_num_reds<-prob_by_num_reds %>% mutate(predicted_prob=itermap_dbl(.x=num_reds, function(.x) sum(num_reds_in_simulation==.x))/num_trials)
print(prob_by_num_reds)
```

### 2.1 (Q7)

```{r}
prob_by_num_reds %>% rename(TheoreticalProbability=prob,EstimatedProbability=predicted_prob) %>% ggplot() + geom_line(aes(x=num_reds, y=TheoreticalProbability)) + geom_line(aes(x=num_reds, y=EstimatedProbability), linetype='dashed') + geom_point(aes(x=num_reds, y=EstimatedProbability)) + theme_bw() + xlab("Number of reds") + ylab("Probabilities")

```

## 2.2 Sampling without replacement

### 2.2 (Q1)

```{r}

set.seed(66)
num_trials <- 1000
sample_size <- 10

bag <- c(rep("red", 50), rep("blue", 30), rep("green", 20))

perform_trial <- function() {
  sample <- sample(bag, sample_size, replace = FALSE)
  
  red_count <- sum(sample == "red")
  blue_count <- sum(sample == "blue")
  green_count <- sum(sample == "green")

  return(list(red = red_count, blue = blue_count, green = green_count))
}

samples <- itermap(1:num_trials, function(x) perform_trial())

min_colors <- itermap_dbl(samples, function(sample_counts) {

  red <- sample_counts$red
  blue <- sample_counts$blue
  green <- sample_counts$green
  return(min(red, blue, green))
})

missing_color_trials <- sum(min_colors == 0)
probability_missing_colors <- missing_color_trials / num_trials

cat("Probability that one or more colors are missing:", probability_missing_colors, "\n")
```

### 2.2 (Q2)

```{r}
all<-choose(100,10)
miss1<-choose(50,10)+choose(30,10)+choose(20,10)
miss2<-choose(80,10)+choose(70,10)+choose(50,10)
p=(miss1+miss2)/all
print(p)
```
In the case of a random seed of 66, the deviation of 1000 simulations is about 2.3%.




















